Quantum Interval-Valued Probability: Contextuality and the Born Rule

### 2017-09-13 prop_letter.tex
prop_letter.tex \section{Gleason} Updating...

The quantum probability measure~$\mu_{1}$ maps some projectors to
$0$ or $0.5$, which are replaced by $\left[0,0\right]$ or $\left[0.5,0.5\right]$
to get $\bar{\mu}_{3}$. This is replacing two numbers by the smallest
intervals containing them. In general, any set of the numbers can
be replaced by the interval containing them if we carefully pick a
set of overlapping intervals. For example, consider the following
collection of intervals: 
\begin{eqnarray}
\widehat{\mathscr{I}}_{n} & = & \set{\left[\frac{k-1}{2n},\frac{k+1}{2n}\right]\cap\left[0,1\right]}{k=0,\ldots,2n}\nonumber \\
 &  & \cup\{\imposs,\necess\}
\end{eqnarray}
which are overlapping intervals with length~$1/n$ except the intervals
at length~$\left(2n\right)^{-1}$ at each end. Then, any Born rule
probability~$\mu_{\rho}^{\mathrm{B}}$ on a three-dimensional Hilbert
space can induce a QIVPM over the intervals above as follows. We first
define a map $\widehat{\iota}_{n}$ from $[0,1]$ to $\widehat{\mathscr{I}}_{n}$
as: 
\begin{itemize}
\item when $x=0$ or $x=1$, $\widehat{\iota}_{n}\left(x\right)=\left[x,x\right]$; 
\item when $0<x<1$, $\widehat{\iota}_{n}\left(x\right)=\left[\frac{k-1}{2n},\frac{k+1}{2n}\right]\cap\left[0,1\right]$,
where $k$ is defined by one of the following: 
\begin{enumerate}
\item $k$ even and $\frac{k-0.5}{2n}\le x\le\frac{k+0.5}{2n}$; 
\item $k$ odd and $\frac{k-0.5}{2n}<x<\frac{k+0.5}{2n}$. 
\end{enumerate}
\end{itemize}
It is now not difficult to verify that $\widehat{\iota}_{n}\circ\mu_{\rho}^{\mathrm{B}}$
is a QIVPM. For example, consider $n=2$, 
\begin{eqnarray}
\widehat{\mathscr{I}}_{1} & = & \set{\left[\frac{k-1}{4},\frac{k+1}{4}\right]\cap\left[0,1\right]}{k=0,\ldots,4}\nonumber \\
 &  & \cup\{\imposs,\necess\}\nonumber \\
 & = & \left\{ \imposs,\left[0,\frac{1}{4}\right],\left[0,\frac{1}{2}\right],\left[\frac{1}{4},\frac{3}{4}\right],\right.\nonumber \\
 &  & \left.\left[\frac{1}{2},1\right],\left[\frac{1}{4},1\right],\necess\right\} \,.
\end{eqnarray}
Then, $\widehat{\iota}_{2}\circ\mu_{1}$, the QIVPM induced by $\mu_{1}$,
can be computed by Table~\ref{tab:Compute-the-QIVPM}.
\begin{table}
\caption{\label{tab:Compute-the-QIVPM}Compute the QIVPM~$\widehat{\iota}_{2}\circ\mu_{1}$
from $\mu_{1}$}
\end{table}

>>>>

The quantum probability measure~$\mu_{1}$ maps some projectors to
$0$ or $0.5$, which are replaced by $\left[0,0\right]$ or $\left[0.5,0.5\right]$
to get $\bar{\mu}_{3}$. This is replacing two numbers by the smallest
intervals containing them. In general, any set of the numbers can
be replaced by the interval containing them if we carefully pick a
set of overlapping intervals. For example, consider the following
collection of intervals: 
\begin{eqnarray}
\widehat{\mathscr{I}}_{n} & = & \set{\left[\frac{k-1}{2n},\frac{k+1}{2n}\right]\cap\left[0,1\right]}{k=0,\ldots,2n}\nonumber \\
 &  & \cup\{\imposs,\necess\}
\end{eqnarray}
which are overlapping intervals with length~$1/n$ except the intervals
at length~$\left(2n\right)^{-1}$ at each end. Then, we can define
a QIVPM mapping to $\widehat{\mathscr{I}}_{n}$ based on any Born
rule probability~$\mu_{\rho}^{\mathrm{B}}$ on a three-dimensional
Hilbert space by replacing the probability~$\mu_{\rho}^{\mathrm{B}}\left(P\right)$
with the interval containing it, $\widehat{\iota}_{n}\left(\mu_{\rho}^{\mathrm{B}}\left(P\right)\right)$.
Since intervals in $\widehat{\mathscr{I}}_{n}$ are overlapping, if
$\mu_{\rho}^{\mathrm{B}}\left(P\right)$ are contained in both intervals~$[\ell_{1},r_{1}]$
and $[\ell_{2},r_{2}]$, $\widehat{\iota}_{n}\left(\mu_{\rho}^{\mathrm{B}}\left(P\right)\right)$
will be $[\ell_{i},r_{i}]$ if $\widehat{\iota}_{n}\left(\mu_{\rho}^{\mathrm{B}}\left(P\right)\right)$
is closer to the center of $[\ell_{i},r_{i}]$. When there is a tie,
we set $\widehat{\iota}_{n}\left(\mu_{\rho}^{\mathrm{B}}\left(P\right)\right)=\left[\frac{k-1}{2n},\frac{k+1}{2n}\right]\cap\left[0,1\right]$
if $k$ even. For example, consider $n=2$, 
\begin{eqnarray}
\widehat{\mathscr{I}}_{1} & = & \set{\left[\frac{k-1}{4},\frac{k+1}{4}\right]\cap\left[0,1\right]}{k=0,\ldots,4}\nonumber \\
 &  & \cup\{\imposs,\necess\}\nonumber \\
 & = & \left\{ \imposs,\left[0,\frac{1}{4}\right],\left[0,\frac{1}{2}\right],\left[\frac{1}{4},\frac{3}{4}\right],\right.\nonumber \\
 &  & \left.\left[\frac{1}{2},1\right],\left[\frac{1}{4},1\right],\necess\right\} \,.
\end{eqnarray}
$\widehat{\iota}_{2}\circ\mu_{1}$ can be computed by Table~\ref{tab:Compute-the-QIVPM}.
\begin{table}
\caption{\label{tab:Compute-the-QIVPM}Compute the QIVPM~$\widehat{\iota}_{2}\circ\mu_{1}$
from $\mu_{1}$ on the minimal subspace of events containing $\proj{0}$,
$\proj{1}$, and $\proj{2}$.}
\begin{tabular}{ccc}
\toprule 
$P$ & $\mu_{1}\left(P\right)$ & $\widehat{\iota}_{2}\left(\mu_{1}\left(P\right)\right)$\tabularnewline
\midrule
\midrule 
$\mathbb{0}$ & $0$ & $\left[0,0\right]$\tabularnewline
\midrule 
$\proj{0}$ & $0.5$ & $\left[\frac{1}{4},\frac{3}{4}\right]$\tabularnewline
\midrule 
$\proj{1}$ & $0.5$ & $\left[\frac{1}{4},\frac{3}{4}\right]$\tabularnewline
\midrule 
$\proj{2}$ & $0$ & $\left[0,0\right]$\tabularnewline
\midrule 
$\mathbb{1}-\proj{0}$ & $0.5$ & $\left[\frac{1}{4},\frac{3}{4}\right]$\tabularnewline
\midrule 
$\mathbb{1}-\proj{1}$ & $0.5$ & $\left[\frac{1}{4},\frac{3}{4}\right]$\tabularnewline
\midrule 
$\mathbb{1}-\proj{2}$ & $1$ & $\left[1,1\right]$\tabularnewline
\midrule 
$\mathbb{1}$ & $1$ & $\left[1,1\right]$\tabularnewline
\bottomrule
\end{tabular}
\end{table}
Let $\bar{\mu}_{4}$ be $\widehat{\iota}_{2}\circ\mu_{1}$. It is
clear that $\bar{\mu}_{4}$ satisfies Eqs.~(\ref{eq:QIVPM-constraints}).
For Eq.~(\ref{eq:QuantumInterval-valuedProbability-Inclusion}),
given a pair of commuting projectors~$P_{0}$ and $P_{1}$ on a three-dimensional
Hilbert space, one of the following must hold:
\begin{enumerate}
\item When the pair of $P_{0}$ and $P_{1}$ is the same as the pair $P_{0}P_{1}$
and $P_{0}+P_{1}-P_{0}P_{1}$, we have $\bar{\mu}_{4}\left(P_{0}+P_{1}-P_{0}P_{1}\right)+\bar{\mu}_{4}\left(P_{0}P_{1}\right)=\bar{\mu}_{4}\left(P_{0}\right)+\bar{\mu}_{4}\left(P_{1}\right)$,
so that Eq.~(\ref{eq:QuantumInterval-valuedProbability-Inclusion})
clearly holds.
\item When either $P_{0}P_{1}=\mathbb{0}$ or $P_{0}+P_{1}-P_{0}P_{1}=\mathbb{1}$,
we must have 
\begin{eqnarray}
 &  & \bar{\mu}_{4}\left(P_{0}+P_{1}-P_{0}P_{1}\right)+\bar{\mu}_{4}\left(P_{0}P_{1}\right)\nonumber \\
 & \subseteq & \left[\mu_{1}\left(P_{0}+P_{1}-P_{0}P_{1}\right)+\mu_{1}\left(P_{0}P_{1}\right)-\frac{1}{4},\right.\nonumber \\
 &  & \left.\mu_{1}\left(P_{0}+P_{1}-P_{0}P_{1}\right)+\mu_{1}\left(P_{0}P_{1}\right)+\frac{1}{4}\right]\nonumber \\
 & = & \left[\mu_{1}\left(P_{0}\right)+\mu_{1}\left(P_{1}\right)-\frac{1}{4},\mu_{1}\left(P_{0}\right)+\mu_{1}\left(P_{1}\right)+\frac{1}{4}\right]\nonumber \\
 & \subseteq & \bar{\mu}_{4}\left(P_{0}\right)+\bar{\mu}_{4}\left(P_{1}\right)\,.
\end{eqnarray}
\end{enumerate}
Therefore, $\bar{\mu}_{4}$ is a QIVPM.
